
```{r settings_1, include=FALSE}
#install needed packages more easily with pacman
if (!require("pacman")) install.packages("pacman")

pacman::p_load(                                # Install or/and import libraries.
               "ggplot2",
               "tidyverse",
               "dplyr",
               "wbstats",
               "gmm",
               "reticulate",
               "data.table",
               "panelvar",
               "pdynmc",
               "tidyr",
               "plm" # packages
)


Sys.setlocale("LC_TIME", "English")

```

# Import  dataset from world databank

```{r}
# Define indicators
my_indicators <- c(                         
  C = "EN.ATM.CO2E.PC",  #CO2 emissions (metric tons per capita)
  E ="EG.USE.PCAP.KG.OE",  #Energy use (kg of oil equivalent per capita)
  Y = "NY.GDP.PCAP.KD", # GDP per capita (constant 2015 US$)
  URB = "SP.URB.TOTL.IN.ZS", #Urban population (% of total population)
  TR = "NE.TRD.GNFS.ZS" # Trade (% of GDP)
)

# Query data
world_stats <- wb_data(my_indicators, start_date = 1990, end_date = 2012) 

```


# Extract "country" "date"    "E"       "C"       "TR"      "Y"       "URB" columns
```{r}
df <- world_stats[,3:length(names(world_stats))]
head(df)
```

# structure of dataset

```{r}
glimpse(df)
```
 
# Missing variables check

```{r}
print(colSums(is.na(df)))
```

# Handle Missing values
In article missing values haven't been metioned that's why we will use just complete case, namely
dataset without NAs

```{r}
df <- df[complete.cases(df),]  #only complete dataset
```

# Create Squared per capita (quadric form)

```{r}
#Create Y^2
df$logY2 = log(df$Y)*log(df$Y)
```

# checking Missing Variables

```{r}
print(colSums(is.na(df)))
```


# Create an ID column for Panel data analysis based on the countries

```{r}
df <- data.table(df)
df[, id := .GRP, by = country]
```
# Split dataset to  European and North Asian region Countries


```{r}
EuropeAsia = c("Albania", "Belgium", "Bulgaria", "Denmark", "France",
"Germany", "Greece", "Hong Kong SAR, China", "Korea, Rep.", "Hungary", "Iceland", "Ireland",
"Italy", "Japan", "Luxembourg", "Netherlands", "Norway", "Portugal",
"Spain", "Sweden", "Switzerland", "United Kingdom")

print("############### European and North Asian region Countries: #########################")
dfEuropeAsia <- df[df$country %in% EuropeAsia,]

sort(unique(dfEuropeAsia$country))
print("#########################################################################################")
paste0("Number of European and North Asian region countries: ",length(unique(dfEuropeAsia$country)))

```

Please note that "Hong Kong SAR, China"  dropped from our reproducible research
as it CO2 of "Hong Kong SAR, China" doesn't have any values.

```{r}
print(nrow(world_stats[world_stats$country == "Hong Kong SAR, China",]))
print(colSums(is.na(world_stats[world_stats$country == "Hong Kong SAR, China",])))
```

# Split dataset to  Latin American and Caribbean region Countries

```{r}
LACarrb = c("Argentina", "Bolivia", "Brazil", "Nicaragua", "Chile",
"Costa Rica", "Ecuador", "Guatemala", "Honduras", "Mexico", "Panama",
"Paraguay", "Peru", "Uruguay", "Venezuela, RB")

print("############### Latin American and Caribbean region Countries: #########################")
dfLACarrb <- df[df$country %in% LACarrb,]

sort(unique(dfLACarrb$country))
print("#########################################################################################")
paste0("Number of Latin American and Caribbean region: ",length(unique(dfLACarrb$country)))

```

Please note that "Venezuela, RB"  dropped from our reproducible research
as it Y of "Venezuela, RB" doesn't have any values.

```{r}
print(nrow(world_stats[world_stats$country == "Venezuela, RB",]))
print(colSums(is.na(world_stats[world_stats$country == "Venezuela, RB",])))
```

# Split dataset to  Middle Eastern, North African, and sub-Saharan region Countries

```{r}

MeNoAfrSub= c("Algeria", "Botswana",
"Cameroon", "Congo, Dem. Rep.", "Cote d'Ivoire", "Ethiopia", "Gabon", "Ghana",
"Egypt, Arab Rep.", "Iran, Islamic Rep.", "Jordan", "Kenya", "Morocco", "Mozambique", "South Africa",
"Senegal", "Sudan", "Syrian Arab Republic", "Togo", "Tunisia","Zambia")

print("############### Middle Eastern, North African, and sub-Saharan region Countries: #########################")
dfMeNoAfrSub <- df[df$country %in% MeNoAfrSub,]

sort(unique(dfMeNoAfrSub$country))
print("#########################################################################################")
paste0("Number of Middle Eastern, North African, and sub-Saharan region: ",length(unique(dfMeNoAfrSub$country)))

```

# Create Global Panel Countries

```{r}

globalPanel= c(EuropeAsia,LACarrb,MeNoAfrSub)

print("############### Global Panel Countries: ################################################")
dfglobalPanel <- df[df$country %in% globalPanel,]

sort(unique(dfglobalPanel$country))
print("#########################################################################################")
paste0("Number of Global Panel: ",length(unique(dfglobalPanel$country)))

```

# Output of Table 2 page: 1105

```{r}
library(pastecs)

statsdfEuropeAsia = round(stat.desc(dfEuropeAsia[,c("E","Y","C","URB","TR")]),3) 

statsdfEuropeAsiaTable2 = statsdfEuropeAsia[c("mean", "std.dev", "coef.var"),]

print(statsdfEuropeAsiaTable2)

```
# Output of Table 2 page: 1105

```{r}
library(pastecs)

statsdfLACarrb = round(stat.desc(dfLACarrb[,c("E","Y","C","URB","TR")]),3) 

statsdfLACarrbTable2 = statsdfLACarrb[c("mean", "std.dev", "coef.var"),]

print(statsdfLACarrbTable2)

```
# Output of Table 2 page: 1105

```{r}
library(pastecs)

statsdfMeNoAfrSub = round(stat.desc(dfMeNoAfrSub[,c("E","Y","C","URB","TR")]),3) 

statsdfMeNoAfrSubTable2 = statsdfMeNoAfrSub[c("mean", "std.dev", "coef.var"),]

print(statsdfMeNoAfrSubTable2)

```

# Table 3 page 1106

```{r}
dflogCorr <- dfglobalPanel[, c("C","Y","E")]
dflogCorr[] <- lapply(dflogCorr, log)
dflogCorr[, c("URB","TR")] <- dfglobalPanel[, c("URB","TR")]
corrMatrix <- cor(dflogCorr)
corrMatrix
```

```{r}
# Hide upper triangle
upper<-corrMatrix
upper[upper.tri(corrMatrix)]<-""
upper<-as.data.frame(upper)
upper
```


