---
title: "A Project for Reproducible Research"
subtitle: "PDF and LaTeX"
author: 
  - Adnan Sevinc (437971) 
  - Le Nhat Tung (426246)
  - Temmuz Burak Yavuzer (444130)
date: today
format: 
  pdf: 
    citation_package: natbib
  html:
    toc: true
    toc-depth: 2
    toc-location: left
    smooth-scroll: true
documentclass: article
number-sections: true
number-depth: 5
title-block-banner: true
fontsize: 11pt
geometry: margin=1in
latex-tinytex: false
output: html_document
---


------------------------------------------------------------------------

**This material may not be copied or published elsewhere (including Facebook and other social media) without the  permission of the authors!**

# Introduction

::: {.justify}

The aim of this project is to assess the validity of the published [article](https://www.sciencedirect.com/science/article/abs/pii/S1364032116000848)
from a prominent publisher. namely, to check whether the published article is
consistent or not.  

:::
     
::: {.justify}

We decided to employ the reproducible research on the selected article
because of the fact that there are many articles in the literature in order to
investigate the relationship between energy consumption and energy use and
their result is somehow contradictory. The relationship between energy
consumption and carbon dioxide (indicator of environmental deterioration)
observed differently due to either dataset or/and region. Considering all of the
inconsistency it will be a great experience for us to check its validity.  

:::


The dataset of the article obtained from [World Data Bank Indicator](http://data.worldbank.org/indicator)   

The code and report of our project can be accessed from [repository](https://github.com/lenhattung/RRProject_426246_437971_444130.git)

------------------------------------------------------------------------

# Install Required Packages

```{r}
if (!require("pacman")) install.packages("pacman")

pacman::p_load(                                
               "ggplot2",
               "tidyverse",
               "dplyr",
               "wbstats",
               "gmm",
               "reticulate",
               "data.table",
               "panelvar",
               "pdynmc",
               "tidyr",
               "plm",
               "pastecs",
               "stargazer",
               "scales"
)


Sys.setlocale("LC_TIME", "English")

```

------------------------------------------------------------------------

# Dataset

Following variables used during modeling process for the time period spanning from 1990 to 2012.

| Variables                                    | Abbreviation |
|----------------------------------------------|--------------|
| CO2 emissions (metric tons per capita)       | C            |
| Energy use (kg of oil equivalent per capita) | E            |
| GDP per capita (constant 2015 US$)           | Y            |
| Urban population (% of total population)     | URB          |
| Trade (% of GDP)                             | TR           |



------------------------------------------------------------------------

# Import Dataset

We imported dataset directly from WDI using **wbstats** packages.

```{r}
my_indicators <- c(                         
  C = "EN.ATM.CO2E.PC",  
  E ="EG.USE.PCAP.KG.OE",  
  Y = "NY.GDP.PCAP.KD", 
  URB = "SP.URB.TOTL.IN.ZS", 
  TR = "NE.TRD.GNFS.ZS" 
)

world_stats <- wb_data(my_indicators, start_date = 1990, end_date = 2012) 
```

C, E, Y, URB, and TR represent CO2 emissions, energy use, GDP per capita, urban population, and trade, respectively.


------------------------------------------------------------------------

# Data Preparation

```{r}
names(world_stats)
```
The output above shows that the iso2c and iso3c columns have been imported into the data environment. Nevertheless, as they are not essential for the analysis, we will eliminate them.

##   Discard the irrelevant columns

```{r}
#| echo: false
df <- world_stats[,3:length(names(world_stats))]
```

iso2c and iso3c columns eliminated from the model.

```{r}
#| echo: false
cat("The dataset consist of",length(setdiff(names(df),c("country","date"))),"variables and",nrow(df), "observations.")
```

We used the following variables to construct a panel data model

```{r}
#| echo: false
print(setdiff(names(df),c("country","date")))
```

##   Completeness check


```{r}
#| echo: false
print(colSums(is.na(select(df,c(E,C,TR,Y,URB)))))
```

As it can be observed, we do have missing values in our dataset.

The article does not mention handling missing values; therefore, we will use complete cases, specifically the dataset without any missing values (NAs).

```{r}
#| echo: false
df <- df[complete.cases(df),]  
```

##    Create a Variable

To construct a model for the Environmental Kuznets Curve (EKC), we generated a squared per capita variable.

We will create $y^2$

```{r}
df$logY2 = log(df$Y)*log(df$Y)
```

## Create an ID column for Panel data analysis based on the countries

For the panel data analysis, we generated an identifier based on the countries.

```{r}
df <- data.table(df)
df[, id := .GRP, by = country]
```

## Split dataset to  based on the regions

We divided the dataset into European and North Asian region countries to build a model that incorporates these regions along with the entire panel data.    

These countries has been indicated in the article.


###   European and North Asian region Countries

```{r}
#| echo: false
EuropeAsia = c("Albania", "Belgium", "Bulgaria", "Denmark", "France",
"Germany", "Greece", "Hong Kong SAR, China", "Korea, Rep.", "Hungary", "Iceland", "Ireland",
"Italy", "Japan", "Luxembourg", "Netherlands", "Norway", "Portugal",
"Spain", "Sweden", "Switzerland", "United Kingdom")

dfEuropeAsia <- df[df$country %in% EuropeAsia,]

sort(unique(dfEuropeAsia$country))
```

```{r}
#| echo: false
paste0("Number of European and North Asian region countries: ",length(unique(dfEuropeAsia$country)))
```

It is worth mentioning that "Hong Kong SAR, China," was included in the author's analysis, but we could not use it in our replicable research because there was a total absence of CO2 emission data, with all values missing.

###   Latin American and Caribbean region Countries

```{r}
#| echo: false
LACarrb = c("Argentina", "Bolivia", "Brazil", "Nicaragua", "Chile",
"Costa Rica", "Ecuador", "Guatemala", "Honduras", "Mexico", "Panama",
"Paraguay", "Peru", "Uruguay", "Venezuela, RB")

dfLACarrb <- df[df$country %in% LACarrb,]

sort(unique(dfLACarrb$country))

paste0("Number of Latin American and Caribbean region: ",length(unique(dfLACarrb$country)))

```

It is worth mentioning that "Venezuela, RB" was included in the author's analysis, but we could not use it in our replicable research because there was a total absence of GDP per capita data, with all values missing.


###   Middle Eastern, North African, and sub-Saharan region Countries

```{r}
#| echo: false
MeNoAfrSub= c("Algeria", "Botswana",
"Cameroon", "Congo, Dem. Rep.", "Cote d'Ivoire", "Ethiopia", "Gabon", "Ghana",
"Egypt, Arab Rep.", "Iran, Islamic Rep.", "Jordan", "Kenya", "Morocco", "Mozambique", "South Africa",
"Senegal", "Sudan", "Syrian Arab Republic", "Togo", "Tunisia","Zambia")

dfMeNoAfrSub <- df[df$country %in% MeNoAfrSub,]

sort(unique(dfMeNoAfrSub$country))

paste0("Number of Middle Eastern, North African, and sub-Saharan region: ",length(unique(dfMeNoAfrSub$country)))

```

###   Creating a Global Panel Countries

```{r}
#| echo: false
globalPanel= c(EuropeAsia,LACarrb,MeNoAfrSub)

dfglobalPanel <- df[df$country %in% globalPanel,]

sort(unique(dfglobalPanel$country))

paste0("Number of Global Panel: ",length(unique(dfglobalPanel$country)))

```

------------------------------------------------------------------------

# Reproducible Outputs - Exploratory Data Analysis

##    Output of Table 2; page 1105 in article

```{r}
#| echo: false
library(pastecs)
statsdfEuropeAsia = round(stat.desc(dfEuropeAsia[,c("E","Y","C","URB","TR")]),3) 
statsdfEuropeAsiaTable2 = statsdfEuropeAsia[c("mean", "std.dev", "coef.var"),]
print(statsdfEuropeAsiaTable2)

```

The coefficient of variation measures the relative variability of a variable compared to its mean. A higher coefficient of variation indicates greater relative variability. In this case, our variables GDP per capita (Y) and Trade (TR) have the highest coefficient of variation for European and North Asian region Countries, suggesting they exhibit relatively more variability compared to their means.



##    Output of Table 2 ( page 1105 in article )

```{r}
#| echo: false
library(pastecs)
statsdfLACarrb = round(stat.desc(dfLACarrb[,c("E","Y","C","URB","TR")]),3) 
statsdfLACarrbTable2 = statsdfLACarrb[c("mean", "std.dev", "coef.var"),]
print(statsdfLACarrbTable2)
```

The mean CO2 emissions per capita (C) in the dataset is 1.764 metric tons for Latin American and Caribbean region Countries, while the mean energy use per capita (E) is 899.041 kg of oil equivalent. The coefficient of variation measures the relative variability of the data, and it ranges from 0.480 to 0.617 across the variables.

##    Output of Table 2 ( page 1105 in the article )

```{r}
#| echo: false
library(pastecs)
statsdfMeNoAfrSub = round(stat.desc(dfMeNoAfrSub[,c("E","Y","C","URB","TR")]),3) 
statsdfMeNoAfrSubTable2 = statsdfMeNoAfrSub[c("mean", "std.dev", "coef.var"),]
print(statsdfMeNoAfrSubTable2)
```
The mean energy use per capita (E) in the dataset is 807.440 kg of oil equivalent for Middle Eastern, North African, and sub-Saharan region Countries , while the mean GDP per capita (Y) is 2442.285 constant 2015 US dollars. The coefficient of variation, which measures the relative variability of the data, ranges from 0.778 to 1.107 across the variables


From the comparison, we can observe variations in the statistical measures across the datasets:

*  The Europe-Asia dataset generally exhibits higher mean values for variables such as energy use (E) and GDP per capita (Y) compared to the other two datasets.
*  The MeNoAfrSub dataset shows the highest standard deviations for all variables, indicating greater variability in the data compared to the other datasets.
*  The coefficient of variation, which measures the relative variability, is highest for the MeNoAfrSub dataset in most cases, suggesting relatively higher dispersion of data compared to the other datasets.


##    Output of Table 3 ( page 1106 in the article )

```{r}
#| echo: false
dflogCorr <- dfglobalPanel[, c("C","Y","E")]
dflogCorr[] <- lapply(dflogCorr, log)
dflogCorr[, c("URB","TR")] <- dfglobalPanel[, c("URB","TR")]
corrMatrix <- cor(dflogCorr)
corrMatrix
```
Based on the correlation matrix, we can observe the following:

*  CO2 emissions (C) and GDP per capita (Y) have a strong positive correlation.
*  CO2 emissions (C) and Energy use (E) have a strong positive correlation.
*  GDP per capita (Y) and Energy use (E) have a strong positive correlation.
*  Urban population (URB) has a moderate positive correlation with CO2 emissions (C), GDP per capita (Y), and Energy use (E).
*  Trade (TR) has a weak positive correlation with CO2 emissions (C), GDP per capita (Y), and Energy use (E).
These correlations provide insights into the relationships between the variables in the dataset, indicating which variables tend to move together or have an impact on each other.


```{r}
#| echo: false
upper<-corrMatrix
upper[upper.tri(corrMatrix)]<-""
upper<-as.data.frame(upper)
upper
```
The upper triangular matrix helps visualize the correlations without duplication. Also matrix provides a compact representation of the correlations, showing the non-redundant information. It is helpful for focusing on the relationships between variables without repeating the same correlations.

------------------------------------------------------------------------

# Graphs of European and Asian Countries for Exploratory Data Analysis

In this section, we will depict the relationship between important variables only for European and Asian countries.

::: {.panel-tabset}

## Energy Use vs. CO2 Emissions

```{r}
#| echo: false
#| warning: false
ggplot(dfEuropeAsia, aes(x = E, y = C)) +
  geom_point(size = 2, alpha = 0.6, color = "#1f77b4") +
  labs(x = "CO2 Emissions (metric tons per capita)",
       y = "Energy Use (kg of oil equivalent per capita)",
       title = "Energy Use vs. CO2 Emissions for Europe & Asia") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 10),
        legend.position = "none") +
  scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = comma)

```

*   The majority of the population in the countries of Europe and Asia have concentrated energy use from 5 to 12.5 kg of oil equivalent per capita. The CO2 emission is equivalent to about 2500 to 7500 tons per capita.    

*   We noticed a positive correlation between energy use and CO2 emissions until the CO2 emissions reached 10,000. However, after reaching this point, the relationship no longer exhibits an upward trend.

## Urban Population vs CO2 Emissions

```{r}
#| echo: false
#| warning: false
ggplot(dfEuropeAsia, aes(x = URB, y = C)) +
  geom_point(size = 2, alpha = 0.6, color = "#1f77b4") +
  labs(x = "Urban Population (% of total population)",
       y = "CO2 Emissions (metric tons per capita)",
       title = "Urban Population vs CO2 Emissions for Europe & Asia") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 10),
        legend.position = "none") +
  scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = comma)

```

*   With the percentage of population in urban areas below 55%, CO2 consumption is quite low. Most cities have CO2 emissions of between 5 and more than 10 tons per capita. There are also special situations where a few municipalities with a population distribution of 80 to 90% have extremely large CO2 emissions, about 20 to 30 tons per capita.

## Average GDP per Capita

```{r}
#| echo: false
#| warning: false
avg_gdp <- aggregate(Y ~ country, dfEuropeAsia, mean)

ggplot(avg_gdp, aes(x = country, y = Y)) +
  geom_bar(stat = "identity", fill = "#4e79a7", width = 0.7) +
  labs(x = "Country", y = "GDP per Capita", title = "Average GDP per Capita for Europe & Asia") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(size = 18, face = "bold", margin = margin(b = 20)),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    legend.position = "none"
  )

```

*    It is evident that Luxembourg has the highest GDP per capita among the other countries, followed by Switzerland, whereas Albania has the lowest average GDP per capita.


:::


------------------------------------------------------------------------

# Dynamic Panel Data Model Estimation

##    Arellano and Bond (1991) twostep estimation extended by nonlinear moment (GMM)

The below hypothesis has been seek by research for e regions together with global panel

**Hypothesis**:  

if $Y$ (GDP per capita) is positive and $Y^2$ (Squared GDP per capita) is negative we do have inverted U-shape pattern   

The following tests should be satisfied for the Generalized Method of Moments (GMM) estimation   

**Sargan Test**:    

$H_o$: Overidentifying restrictions valid    

In this context, we are looking for a p-value greater than 0.05.    
For GMM estimation overidentifying restrictions have to be valid

**Autocorrelation test (2)**:   

$H_o$: no serial correlation of order 2 in epsilon

In this context, we are looking for a p-value greater than 0.05.            
For GMM estimation we should not have second order serial correlation

###   Output of Table 4 (page 1107 in the article) : Result for the global panel


```{r}
#| echo: false
#| warning: false
globalPanelGMM <- pgmm(log(C) ~ lag(log(C), 1) + log(Y) + logY2 +log(E)  + log(URB) + log(TR)| lag(log(C), 0:20),
           data = dfglobalPanel, effect = "twoways", model = "onestep", 
           transformation = "ld")
```

```{r}
#| echo: false
#| warning: false
summary(globalPanelGMM, robust = TRUE)
```

*   For the global panel the null hypothesis of no serial correlation (Autocorrelation test (2)) cannot be rejected, given the p-value > 0.05 and absence of a second order of correlation.

*   For the global panel the Sargan test statistics to assess the validity of overidentifying restrictions.The test yielded a p-value of 1, and thus we fail to reject the null hypothesis that the overidentifying restrictions are valid.

*   For the global panel The results of the **Sargan** test and **AR(2)** analysis are consistent with the author's findings.

*   In contrast to the author's findings of the global panel , our panel data model shows that all variables are statistically significant, including Trade.

*   Our model of the the global panel indicates a positive effect of lagged CO2 emissions on current CO2 emissions, whereas the author's findings of global panel showed a negative effect of lagged CO2 emissions on current CO2 emissions.

*   Our model of the the global panel indicates a positive impact of urban population on CO2 emissions, whereas the author's findings of global panel showed a negative impact of urban population on CO2 emissions.

*   Our model of the the global panel indicates that the hypothesis is valid for the global panel due to $log(Y)>0\text{ and }log(Y^2) <0$, However, the author's findings show that the hypothesis is invalid.


###   Output of Table 5,6 (page 1107 and 1108 in the article) : European and North Asian region countries

```{r}
#| echo: false
#| warning: false
EuropeAsiaGMM <- pgmm(log(C) ~ lag(log(C), 1) + log(Y) + logY2 +log(E)  + log(URB) + log(TR)| lag(log(C), 0:20),
           data = dfEuropeAsia, effect = "twoways", model = "onestep", 
           transformation = "ld")
```

```{r}
#| echo: false
#| warning: false
summary(EuropeAsiaGMM, robust = TRUE)
```

*   For European and North Asian region countries  the null hypothesis of no serial correlation (Autocorrelation test (2)) cannot be rejected, given the p-value > 0.05 and absence of a second order of correlation.

*   For European and North Asian region countries the Sargan test statistics to assess the validity of overidentifying restrictions.The test yielded a p-value of 1, and thus we fail to reject the null hypothesis that the overidentifying restrictions are valid.

*   For European and North Asian region countries the results of the **Sargan** test and **AR(2)** analysis are consistent with the author's findings.

*   For European and North Asian region countries While the author's findings suggest that only urban population is statistically insignificant, our model indicates that only energy use is statistically insignificant.

*   For European and North Asian region countries it appears that there is a discrepancy between the author's findings and our model. The author's findings indicate that urban population and trade have a negative impact on CO2 emissions, while our model indicates a positive impact.

*   Our findings indicate that the hypothesis is valid for the European and North Asian region countries, as we observed an inverted U-shape pattern, which is consistent with the author's findings.

###   Output of Table 5,6 (page 1107 and 1108 in the article) : Latin American and Caribbean region countries

```{r}
#| echo: false
#| warning: false
LACarrb <- pgmm(log(C) ~ lag(log(C), 1) + log(Y) + logY2 +log(E)  + log(URB) + log(TR)| lag(log(C), 0:20),
           data = dfLACarrb, effect = "individual", model = "onestep", 
           transformation = "ld")

```

```{r}
#| echo: false
#| warning: false
summary(LACarrb, robust = TRUE)
```

*   For Latin American and Caribbean region countries the null hypothesis of no serial correlation (Autocorrelation test (2)) cannot be rejected, given the p-value > 0.05 and absence of a second order of correlation.

*   For Latin American and Caribbean region countries the Sargan test statistics to assess the validity of overidentifying restrictions.The test yielded a p-value of 1, and thus we fail to reject the null hypothesis that the overidentifying restrictions are valid.

*   For Latin American and Caribbean region countries the results of the **Sargan** test and **AR(2)** analysis are consistent with the author's findings

*   Our model for Latin American and Caribbean region countries suggests that the lagged CO2 emissions variable is statistically significant, while it is not the case in the author's findings.

*   Our findings suggest that the hypothesis is invalid for the Latin American and Caribbean region countries,which is consistent with the author's findings.


###   Output of Table 5,6 (page 1107 and 1108 in the article) : Middle Eastern, North African, and sub-Saharan region countries

```{r}
#| echo: false
#| warning: false
MeNoAfrSub <- pgmm(log(C) ~ lag(log(C), 1) + log(Y) + logY2 +log(E)  + log(URB) + log(TR)| lag(log(C), 0:20),
           data = dfMeNoAfrSub, effect = "twoways", model = "onestep", 
           transformation = "ld")

```

```{r}
#| echo: false
#| warning: false
summary(MeNoAfrSub, robust = TRUE)
```
*   For Middle Eastern, North African, and sub-Saharan region countries the null hypothesis of no serial correlation (Autocorrelation test (2)) cannot be rejected, given the p-value > 0.05 and absence of a second order of correlation.

*   For Middle Eastern, North African, and sub-Saharan region countries the Sargan test statistics to assess the validity of overidentifying restrictions.The test yielded a p-value of 1, and thus we fail to reject the null hypothesis that the overidentifying restrictions are valid.

*   For Middle Eastern, North African, and sub-Saharan region countries the results of the **Sargan** test and **AR(2)** analysis are consistent with the author's findings

*  In the Middle Eastern, North African, and sub-Saharan region, the author's findings suggest that only urban population is statistically insignificant, while our model indicates that only trade is statistically insignificant.

*   For Middle Eastern, North African, and sub-Saharan region Countries our model indicates a positive effect of lagged CO2 emissions on current CO2 emissions, whereas the author's findings showed a negative effect of lagged CO2 emissions on current CO2 emissions.

*   For Middle Eastern, North African, and sub-Saharan region countries our model indicates a positive effect of trade on CO2 emissions, whereas the author's findings  showed a negative effect of trade on CO2 emissions. Nevertheless trade is statistically insignificant in our model

*   Our findings suggest that the hypothesis is valid for the Middle Eastern, North African, and sub-Saharan regions,which is consistent with the author's findings.


------------------------------------------------------------------------

# CONCLUSION


*   Our findings showed that energy use`log(E)` has a positive impact on carbon dioxide emissions for all the panels, which is in line with the authors' results

*   Our findings showed that GDP per capita `log(Y)` has a positive, statistically significant impact on carbon dioxide emissions for the Global Panel, European and North Asian, Middle Eastern, North African, and sub-Saharan African regions, except for the Latin American and Caribbean regions, which is in line with the authors' results     

**Does the presence of an inverted U-shaped curve observed ?**

| Region                                       | The author's findings | Our Findings | Does it align with the author's findings ? |
|----------------------------------------------|-----------------------|--------------|--------------------------------------------|
| Global Panel                                 | Not Valid             | Valid        | No                                         |
| Europe and North Asian                       | Valid                 | Valid        | Yes                                        |
| Latin American and Caribbean                 | Not Valid             | Not Valid    | Yes                                        |
| Middle Eastern, North African and Sub-Sharan | Valid                 | Valid        | Yes                                        |
